Marine species do not need visas: the transboundary nature of the world’s exploited marine species

### 2020-07-29 Scripts/Manuscript_Draft.Rmd
All comments addressed

---
title: "Marine species do not need visas: the transboundary nature of the world’s exploited marine species"
author: "Authors: Juliano Palacios-Abrantes*^1^, Gabriel Reygondeau^1,2^, Colette Wabnitz^1,3^ and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Yale University, New Haven, Connecticut, USA. ^3^Stockholm Resilience Center, Stockholm University, Stockholm, Sweden."
subtitle: "CORECTIONS ARE IN BOLD. DO NOT UNBOLD"
# subtitle: "Most neighbor nations share marine exploited species"
output:
  word_document:
    # reference_docx: ./Manuscript_Instructions/science_template.docx
  pdf_document:
    fig_caption: yes
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ../References/FishForVisa_Ref.bib
csl: ../References/nature_cc.csl
---

@@

## Questions for Gabs:
- Occurence data time frame
- ESM is 1970 to 2000, is this beacuse of cliamte period? Reference?
- SAU-SDM is only one file right, there are no historic estimations
- **Rev 1.** Line 132: Why? These last sentence is hard to understand.

## Questions for William:
- I want to reference both Ocean Canada papers to address a comment, is it fair to reference them as *in press*?

@@

<!-- # Reviewers (5): -->
<!-- *Include Names, affiliations, and e-mail addresses of up to five potential reviewers and up to five excluded reviewers.*   -->

<!-- **Note:** Please include potential reviewers you think appropiate to review the manuscript (Or confirm my suggestions) -->

<!-- - **Potential Reviewers (J)**   -->
<!--  -- Claire Armstrong, The Norwegian College of Fishery Science, Tromso, Norway claire.armstrong@uit.no -->
<!--  -- Gakushi Ishimura, Faculty of Agriculture Department of Food Production and Environmental Management, Iwate University, Morioka, Japan; gakugaku@iwate-u.ac.jp -->
<!--  -- Rebecca Selden, Wellesley College, Wellesley, MA, USA; bselden@wellesley.edu -->
<!--  -- Julia Xue, Shanghai Jiao Tong University; juliaxue@ouc.edu.cn -->

<!-- - **Excluded Reviewers (J)**   -->

# Abstract
<!-- *An opening sentence that sets the question that you address and is comprehensible to the general reader, background content specific to this study, results, and a concluding sentence. It should be a single paragraph.* -->

Regulatory boundaries and species distributions are often not aligned, especially in the case of marine species crossing multiple Economic Exclusive Zones (EEZs). This represents a challenge for fisheries management as policies tend to focus at the national level, yet collaborations are needed to maximize long-term ecological, social and economic benefits. Here, we combined species distributions and the spatial delimitation of EEZs to identify the number of commercially exploited marine species that are shared between neighboring nations. We found that 67% of the world's identified fished species are transboundary (n = 633) and were responsible for 80% of the global-EEZs catches (48 million tons) and 78.5% of fisheries revenue (USD 65 billion) in the last decade. Some countries received over 90% of their catch and economic benefits from a few shared resources. Our analysis suggests that fisheries targeting transboundary species appear **more intensilly fished** than those mainly catching non-transboundary species. Our study has direct implications **for fisheries management** of transboundary species, pointing to the urgent need for strengthened international cooperation in the management of these fisheries.

@@

Distributions of marine species around the world are not constrained by human-made boundaries; rather they are shaped by biotic and abiotic factors defined by their evolutionary history [@Hutchinson:1957nt; @Nelson:2016kk]. Thus, a species can have a broader distribution (cosmopolitan) or be endemic [@Nelson:2016kk]. However, fisheries management is predicated on the definition of “stocks” that, in most cases, are delineated by human-made spatial boundaries that often do not correspond to biologically-meaningful population units [@Song:2017iua; @FredstonHermann:2018kp]. The delineation of Economic Exclusive Zones (EEZs) under the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, for instance, extended political boundaries **from 12 nautical miles (territorial see) to 200 nautical miles** across the distribution of many species, creating shared stocks between nations [@UnitedNations:1986tl]. **Shared stocks can be classified in three non-exclusive categories; transboundary, that is stocks that cross the EEZs of two or more coastal states; straddling, that cross neighboring EEZs and the high seas; and highly migratory stocks, that cross non-neighboring EEZs and the high seas (mainly tunas) [@Munro:2004th]. For the purpose of this research we focus on identifying the transboundary nature of marine fished species, that is, those shared by neighboring EEZs, but noting that in some cases these might also be preset in the high seas (e.g. straddling).** Theories and empirical experience have shown that fisheries targeting resources that straddle between political boundaries complicate fisheries management and potentially reduce its effectiveness to achieve stated objectives [@Munro:2002uf; **@Englander:2019dq**]. For instance, climate-driven changes in species distributions have led to conflict between nations due to changes in the proportion of captures [@Munro:2002uf], quota allocation, and fishery newcomers [@Spijkers:2017ij]. Moreover, climate change is likely to exacerbate these types of conflict and presents new challenges for political relations between neighboring countries [@Pinsky:2018cb] and for fisheries management [@Miller:2013iv].

@@

**We identified 633 transboundary exploited species in the world (67.5% of the species analyzed), almost double the previous estimates [@Teh:2015gd].** Between 2005 and 2014, national fleets targeting transboundary species within EEZs **caught** an **annual** average of 47.6 million tons, representing 79.9% of the EEZ-derived catches reconstructed by the *Sea Around us* at the species level, and generating **a yearly average of USD 65,483 million in fishing revenue during the same time period (Fig. 1)**. This is equivalent to 78.5% of the global fishing revenue of the same period. These values are more than the 34.2 million tons and USD 30,700 million in fishing revenue from shared stocks previously estimated [@Teh:2015gd]. **This result suggests that the participation of transboundary species in global catch and fishing revenue might have been previously underestimated. However, it is important to mention that the referenced article reports yearly FAO reported catch data for both EEZs and the high seas up to 2006. Thus, the divergence between studies can be influenced by the data source and time period in addition to the higher amount of transboundary species identified in the present study. However, using the FAO reported data for 2006, the average participation of the identified transboundary species was 61%, still higher than that reported by Teh and Sumaila (2015).**

Based on our assessment (see Methods), 590 transboundary species were also identified as "discrete" species, meaning that the same species could be shared by two neighboring EEZs (transboundary) but not with a third neighbor (discrete). For example, Peruvian anchoveta (*Engraulis ringens*) is fished by Peru, Chile, and Ecuador but only transboundary between Peru and Chile [@Cashion:2018cg], and thus discrete for Ecuador (**spatial threshold < 10%**). The **annual average contribution of discrete species (in our previous example would be only considering Equator's anchoveta catch) between 2005 and 2014** was 14.8 million tons representing USD 22,167 million in fishing revenue **over the same time period**. Within these species, 305 were considered exclusively discrete, that is, they could be fished by neighboring nations but were not considered transboundary **according to our methods of determination (See methods)**. These species represented only **an annual average of** 2.8 million tons and 3,691 million USD between 2005 and 2014. These results underscore the importance of transboundary species at the global level.

We group countries according to the United Nations sub-regions and found that transboundary species are particularly economically important for Northern America (mean **yearly** revenue = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (2,827 $\pm$ 2,667 million USD). The estimated revenues from transboundary species in these two regions are significantly different from other regions (one-way ANOVA; *DF*(16,158) = 4.948, *p* < 0.001, $\alpha$ = 0.05) (Fig. 1B). China (USD 6,279 million) in **Eastern** Asia and the USA **In Northern America** (USD 9,605 million) contribute 55% and 82% to the **annual average** revenues. In addition to China and the USA, Russia (USD 6,361 million), Peru (USD 5,211 **million**) and Japan (USD 3,368 **million**) are amongst the top five countries with the most fishing revenue generated from transboundary species in the last decade. These five nations are responsible for 40% of the global fisheries revenue from transboundary species.

@@

![Figure 3. Number of transboundary species by catch category. Actegory A, Increasing catch; Category B, stable catch; Category C, decreasing catch](../Figures/Fig3.png)

@@

Over 130 marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk], from highly migratory species associated to pelagic-oceanic ecosystems like tunas (*Thunnus sp.*), to less mobile reef-associated species like greater amberjack (*Seriola dumerili*), to species found in demersal ecosystems like gilthead seabream (*Sparus aurata*). Species associated to pelagic-oceanic ecosystems are the only group whose EEZ range (i.e., the number of EEZs where the species occur as transboundary) was significantly different from other groups (one-way ANOVA; *DF*(5,597) = 53.82, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 EEZs per species. The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these species have a narrower distribution or are less mobile (Fig. 4A). **This result is not surprising as many of the species with preference for pelagic-oceanic ecosystems categorized as transboundary are actually highly migratory straddling species (species that go from one EEZ to another while crossing the high seas). Thus, the number of non-neighboring EEZs sharing a highly migratory species can be over 100, like the case of Bigeye tuna (Thunnus obesus) (Fig. 4A)**

@@

**Due to their large migration patterns and presence in areas beyond national jurisdiction, many highly migratory species are managed by Regional Fisheries Management Organizations (RFMOs). On the other hand, species with narrower ranges are often managed at the national level due to localized populations [@Payne:2008vm] and the potential boundary-crossing of these species must not be ignored.** The world of fisheries is highly connected through larvae dispersal and so, even centralized populations could be supplying populations thousands of kilometers away [@Ramesh:2019va]. In addition, coral regions often share multiple species, making coordination among countries key to achieve fisheries and conservation goals [@Payne:2008vm]. Many coastal communities, particularly in developing regions, are highly dependent on subsistence fishing for food security and livelihoods [@CisnerosMontemayor:2016gq; @Wabnitz:2018gf]. For example, all of the top five countries fishing transboundary reef and pelagic-oceanic associated species (Fig. **4B**) are among the highest fish consumption countries [@FAO:2018KK]. In these countries, both pelagic and reef fishes contribute substantially to both local food security and livelihood [@Cabral:2018hc].

@@

**The Sea Aorund Us has reconstructed global fisheries catches from 1950 to 2014 identifying commercial marine fish and invertebrates as well as fishing regions [@Zeller2016]. We used the Sea Aorund Us reconstructed species list to determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs. The total number of species analyzed was 938 representing 63.2% of identified species and accounting for 96.52% of the catch identified at the species' level.** To determine the current distribution of exploitable marine species, we used four data sources of species-distributions: (*i*) occurrence data, (*ii*) an Ensemble Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (Table 1). Each source represents a different method of estimating the distribution of a given species and thus, provides a more robust result **than using only one data source**. Only commercial fished species with data from all four sources were included in the analysis.

@@

The ENM-Nereus consists in a multimodel approach based on a Bioclim and a Boosted Regression Tree model [@Thuiller:2009gp], a Maxent model [@Phillips:2006ff], and a Non-Parametric Probabilistic Ecological Niche Model [@Beaugrand:2011fd]. Environmental variables utilized in the models include sea surface temperature, surface pH, surface oxygen concentration, and vertically integrated (0–100 m) net primary production (NPP) [@Asch:2018ca]. Global environmental conditions were average for the period 1970-2000 and averaged for three Earth System Models developed by the Geophysical Fluid Dynamics Laboratory (GFDL- https://www.gfdl.noaa.gov/earth-system-model/), the Institute Pierre Simon Laplace (IPSL- www.icmc.ipsl.fr/), and the Max Planck Institute for Meteorology (MPI- www.mpimet.mpg.de/en/science/models/). See [@Asch:2018ca; @Reygondeau:2019uh] for model details.

@@

**We used the occurrence database, the ENM-Nereus model, and SDM-SAU model to determine the presence of each species within each of the $0.5^{\circ}$ x $0.5^{\circ}$ grid cells. All analyses only considered cases where all databases matched to obtain a more conservative estimate of transboundary species. Moreover, we assumed that a species was only present in a given grid-cell if it was reported in the Sea *Around Us catch* database. Therefore, all species that were not reported as caught in any single year between the reference years (2005 to 2014) in a given grid-cell were dropped. This rule assumes that if a commercial species is projected within any fishing country, such a species would have been fished, and thus reported at some point over the last decadde of data (2005 - 2014), thus validating the models and only selected "current" shared speceies.**

@@

Finally, to have a more robust result and not categorize a species as transboundary based on its presence in a single $0.5^{\circ}$ x $0.5^{\circ}$ grid cell within an EEZ, we computed an *Area Index*. The *Area Index* consist in the proportion of a given species' overarching shared distribution between neighboring EEZs accounted for by each individual EEZ. We classified a species as transboundary if both neighboring EEZs enclosed over 25% of the species joint shared distribution. **While a species that has less than the selected treshold is not considered transboundary in this paper**, this threshold can be lowered for a more relaxed result or increased for a more conservative estimate (Fig. S3).

@@

We estimated the economic contribution in **2019** real USD of transboundary species for each country using global ex-vessel price data [@Tai2017]. The database we draw from includes ex-vessel price derived from multiple sources and a structured interpolation method (e.g. similar countries, species) to fill in data gaps [@Sumaila:2015uc]. **The dataset is harmonized with the SAU catch data to estimate yearly fishing revenue (as ex-vessel price) for all species and EEZs of this study (http://www.seaaroundus.org/data/#/feru). We report the average fishing revenue of 2005 to 2014 derivede from the fishing activity within the global EEZs. Alike the catch data, for this study we did not include revenue from areas beyond national jurisdiction. We used the US Gross Domestic Product Implicit Price Deflator (average of all 2019 quarters) according to the US Saint Louis Federal Reserve Bank (https://fred.stlouisfed.org/searchresults?st=GDP+Implicit+price+deflator) to transform the original 2010 real USD value to 2019 real USD.**

We used the catch data to determine the **catch trend** of each species within an EEZ. Although this method has previously been used to estimate stock status [@Grainger:1996tk], the categories presented here are intended as catch trends, and not fishing status for each species as many environmental and social-economic factors (e.g. temperature, markets, fishing policies, **and fishing effort) affect fishing catch volume **[@Pauly:2013gs; @Branch:2008kk]. We only assessed species within each EEZ for which at least ten years of data were available between the first and last reported catch and with at least five consecutive years of data. Three final categories were drawn for each species depending on the catch volume within each EEZ (e.g. present, maximum and minimum EEZ's historical catch) and the year (e.g. year of maximum historical catch of the species within that EEZ) (Table 2) [@Kleisner:2011wn]. Following this concept, Category A would represent fisheries that are **increasing in catch, Category B would be species that have a constant catch rate, and Category C would be species that have seen a decrease in catch in the last 10 years. Finally, we reported the predominant category over the time period considered.**

@@

Categories <- c("A - Growing Fishery",
                  "B - Stable Fishery",
                  "C - Fishery under 10% of Peak",
                  "No Status")

@@

![Figure S1. Number of transboundary species per EEZ and their contributions to countries’ EEZ catch. The number of species is displayed in the EEZ while catch (in thousand tons) is displayed in the land polygon.](./Figures/FigS1.png)

The number of gridcells in which a species is present (*Area Index*) is expected to directly influence whether or not a species is considered transboundary. Therefore, we tested the sensitivity of our results to the *Area Index* by estimating the number of transboundary species along a gradient of values (Fig. S2). As expected, using a more relaxed value (e.g. the species is present within 10% or more of the EEZ) will result in a greater number of transboundary species (median of 40 EEZs shared per species), while a more restrictive value (e.g., 50%) results in a median estimate of 4. The current analysis utilized a 25% *Area Index* threshold (Fig. S2 - darker histogram).

@@

![Figure S2. Histogram of number of transboundary species using different Area Index threshold values. The darker histogram represents the threshold (25%) used in the present analysis](./Figures/FigS4.png)

>>>>

---
title: "Marine species do not need visas: the transboundary nature of the world’s exploited marine species"
author: "Authors: Juliano Palacios-Abrantes*^1^, Gabriel Reygondeau^1,2^, Colette Wabnitz^1,3^ and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Yale University, New Haven, Connecticut, USA. ^3^Stockholm Resilience Center, Stockholm University, Stockholm, Sweden."
subtitle: "CORECTIONS ARE IN BOLD. DO NOT UNBOLD"
# subtitle: "Most neighbor nations share marine exploited species"
output:
  pdf_document:
    fig_caption: yes
  word_document:
    # reference_docx: ./Manuscript_Instructions/science_template.docx
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ../References/FishForVisa_Ref.bib
csl: ../References/nature_cc.csl
---

@@

## Questions for Gabs:
- Teh et al comment. Do we include another data or just the explanation is fine?
- **Rev 1.** Line 132: Why? These last sentence is hard to understand.

@@

# Abstract

Regulatory boundaries and species distributions are often not aligned, especially in the case of marine species crossing multiple Economic Exclusive Zones (EEZs). This represents a challenge for fisheries management as policies tend to focus at the national level, yet collaborations are needed to maximize long-term ecological, social and economic benefits. Here, we combined species distributions and the spatial delimitation of EEZs to identify the number of commercially exploited marine species that are shared between neighboring nations. We found that 67% of the world's identified fished species are transboundary (n = 633) and were responsible for 80% of the global-EEZs catches (48 million tons) and 78.5% of fisheries revenue (USD 65 billion) in the last decade. Some countries received over 90% of their catch and economic benefits from a few shared resources. Our analysis suggests that fisheries targeting transboundary species appear **more intensively fished** than those mainly catching non-transboundary species. Our study has direct implications **for fisheries management** of transboundary species, pointing to the urgent need for strengthened international cooperation in the management of these fisheries.

@@

Distributions of marine species around the world are not constrained by human-made boundaries; rather they are shaped by biotic and abiotic factors defined by their evolutionary history [@Hutchinson:1957nt; @Nelson:2016kk]. Thus, a species can have a broader distribution (cosmopolitan) or be endemic [@Nelson:2016kk]. However, fisheries management is predicated on the definition of “stocks” that, in most cases, are delineated by human-made spatial boundaries that often do not correspond to biologically-meaningful population units [@Song:2017iua; @FredstonHermann:2018kp]. The delineation of Economic Exclusive Zones (EEZs) under the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, for instance, extended political boundaries **from 12 nautical miles (territorial see) to 200 nautical miles** across the distribution of many species, creating shared stocks between nations [@UnitedNations:1986tl]. **Shared stocks can be classified in three non-exclusive categories; transboundary, that is stocks that cross the EEZs of two or more coastal states; straddling, that cross neighboring EEZs and the high seas; and highly migratory stocks, that cross non-neighboring EEZs and the high seas (mainly tunas) [@Munro:2004th]. For the purpose of this research we focus on identifying the transboundary nature of marine fished species, that is, those shared by neighboring EEZs, noting that in some cases these might also be present in the high seas (e.g. straddling).** Theories and empirical experience have shown that fisheries targeting resources that straddle between political boundaries complicate fisheries management and potentially reduce its effectiveness to achieve stated objectives [@Munro:2002uf; **@Englander:2019dq**]. For instance, climate-driven changes in species distributions have led to conflict between nations due to changes in the proportion of captures [@Munro:2002uf], quota allocation, and fishery newcomers [@Spijkers:2017ij]. Moreover, climate change is likely to exacerbate these types of conflict and presents new challenges for political relations between neighboring countries [@Pinsky:2018cb] and for fisheries management [@Miller:2013iv].

@@

**We identified 633 transboundary exploited species in the world (67.5% of the species analyzed), almost double the previous estimates [@Teh:2015gd].** Between 2005 and 2014, national fleets targeting transboundary species within EEZs **caught** an **annual** average of 47.6 million tons, representing 79.9% of the EEZ-derived catches reconstructed by the *Sea Around us* at the species level, and generating **a yearly average of USD 65,483 million in fishing revenue during the same time period (Fig. 1)**. This is equivalent to 78.5% of the global fishing revenue of the same period. These values are more than the 34.2 million tons and USD 30,700 million in fishing revenue from shared stocks previously estimated [@Teh:2015gd]. **This result suggests that the participation of transboundary species in global catch and fishing revenue might have been previously underestimated. However, it is important to mention that the referenced article reports yearly FAO reported catch data for both EEZs and the high seas up to 2006. Thus, the divergence between studies can be influenced by the data source and time period in addition to the higher amount of transboundary species identified in the present study. ... Using the FAO reported data for 2006, the average participation of the identified transboundary species was 61%, still higher than that reported by Teh and Sumaila (2015).**

Based on our assessment (see Methods), 590 transboundary species were also identified as "discrete" species, meaning that the same species could be shared by two neighboring EEZs (transboundary) but not with a third neighbor (discrete). For example, we found that Peruvian anchoveta (*Engraulis ringens*) is fished by Peru, Chile, and Ecuador but only transboundary between Peru and Chile [@Cashion:2018cg], and thus discrete for Ecuador (**spatial threshold < 10%**). The **annual average contribution of discrete species (in our previous example would be only considering Ecuador's anchoveta catch) between 2005 and 2014** was 14.8 million tons representing USD 22,167 million in fishing revenue **over the same time period**. Within these species, 305 were considered exclusively discrete, that is, they could be fished by neighboring nations but were not considered transboundary **according to our methods of determination (See methods)**. These species represented only **an annual average of** 2.8 million tons and 3,691 million USD between 2005 and 2014. These results underscore the importance of transboundary species at the global level.

We group countries according to the United Nations sub-regions and found that transboundary species are particularly economically important for Northern America (**average yearly** revenue = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (2,827 $\pm$ 2,667 million USD). The estimated revenues from transboundary species in these two regions are significantly different from other regions (one-way ANOVA; *DF*(16,158) = 4.948, *p* < 0.001, $\alpha$ = 0.05) (Fig. 1B). China (USD 6,279 million) in **Eastern** Asia and the USA **in Northern America** (USD 9,605 million) contribute 55% and 82% to the **annual average** revenues. In addition to China and the USA, Russia (USD 6,361 million), Peru (USD 5,211 **million**) and Japan (USD 3,368 **million**) are among the top five countries with the most fishing revenue generated from transboundary species in the last decade. These five nations are responsible for 40% of the global fisheries revenue from transboundary species.

@@

![Figure 3. Number of transboundary species by catch trend and EEZ. Category A, Growing trend; Category B, Stable trend; Category C, Shrinking trend](../Figures/Fig3.png)

@@

Over 130 marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk], from highly migratory species associated to pelagic-oceanic ecosystems like tunas (*Thunnus sp.*), to less mobile reef-associated species like greater amberjack (*Seriola dumerili*), to species found in demersal ecosystems like gilthead seabream (*Sparus aurata*). Species associated to pelagic-oceanic ecosystems are the only group whose EEZ range (i.e., the number of EEZs where the species occur as transboundary) was significantly different from other groups (one-way ANOVA; *DF*(5,597) = 53.82, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 EEZs per species. The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these species have a narrower distribution or are less mobile (Fig. 4A). **[Colette] This result is likely an effect of the broad distribution that many of the species with preference for pelagic-oceanic ecosystems have. Many of these species are actually highly migratory straddling species (species that go from one EEZ to another while crossing the high seas). Thus, the number of non-neighboring EEZs sharing a highly migratory species can be over 100, like the case of Bigeye tuna (Thunnus obesus) (Fig. 4A). Due to their large migration patterns and presence in areas beyond national jurisdiction, many highly migratory species are managed by Regional Fisheries Management Organizations (RFMOs).**

CW. because they are migratory they fall under the management of RFMOs...

@@

**However, this category also includes species that are not considered highly migratory but still shared by numerous neighboring EEZs (e.g. Belone belone). Species with narrower ranges are often managed at the national level due to localized populations [@Payne:2008vm] but the potential boundary-crossing of these species must not be ignored. In addition, the world of fisheries is highly connected through larvae dispersal and so, even centralized populations could be supplying populations thousands of kilometers away [@Ramesh:2019va]. Despite coral reef-related species being among the less shared neighboring EEZs identified in this study (Fig. 4), coral regions, often share multiple species through larvae connectivity [@Schill:2015fh] and adult movement [@Levin:2018in]. In some areas like the Caribbean and the Western Indian Ocean, transboundary marine protected areas have been identified as potential tools for achieving fisheries and conservation goals [@Levin:2018in; Perez:2019dm].** The proper management of coral reef species is very important for many coastal communities, particularly in developing regions, that are highly dependent on subsistence fishing for food security and livelihoods [@CisnerosMontemayor:2016gq; @Wabnitz:2018gf]. For example, all of the top five countries fishing transboundary reef and pelagic-oceanic associated species (Fig. **4B**) are among the highest fish consumption countries [@FAO:2018KK]. In these countries, both pelagic and reef fishes contribute substantially to both local food security and livelihood [@Cabral:2018hc].

@@

**The Sea Around Us has reconstructed global fisheries catches from 1950 to 2014 identifying commercial marine fish and invertebrates as well as fishing regions [@Zeller2016]. We used the Sea Around Us reconstructed species list to determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs. The total number of species analyzed was 938 representing 63.2% of identified species and accounting for 96.52% of the catch identified at the species' level.** To determine the current distribution of exploitable marine species, we used four data sources of species-distributions: (*i*) occurrence data, (*ii*) an Ensemble Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (Table 1). Each source represents a different method of estimating the distribution of a given species and thus, provides a more robust result **than using only one data source**. Only commercial fished species with data from all four sources were included in the analysis.

@@

The ENM-Nereus consists in a multimodel approach based on a Bioclim and a Boosted Regression Tree model [@Thuiller:2009gp], a Maxent model [@Phillips:2006ff], and a Non-Parametric Probabilistic Ecological Niche Model [@Beaugrand:2011fd]. Environmental variables utilized in the models include sea surface temperature, surface pH, surface oxygen concentration, and vertically integrated (0–100 m) net primary production (NPP) [@Asch:2018ca]. Global environmental conditions were average for the **30-year climate normal period of 1970-2000** and averaged for three Earth System Models developed by the Geophysical Fluid Dynamics Laboratory (GFDL- https://www.gfdl.noaa.gov/earth-system-model/), the Institute Pierre Simon Laplace (IPSL- www.icmc.ipsl.fr/), and the Max Planck Institute for Meteorology (MPI- www.mpimet.mpg.de/en/science/models/). See [@Asch:2018ca; @Reygondeau:2019uh] for model details.

@@

**We used the occurrence database, the ENM-Nereus model, and SDM-SAU model to determine the presence of each species within each of the $0.5^{\circ}$ x $0.5^{\circ}$ grid cells. All analyses only considered cases where all databases matched to obtain a more conservative estimate of transboundary species. Moreover, we assumed that a species was only present in a given grid-cell if it was reported in the Sea *Around Us catch* database. Therefore, all species that were not reported as caught in any single year between the reference years (2005 to 2014) in a given grid-cell were dropped. This rule assumes that if a commercial species is projected within any fishing country, such a species would have been fished, and thus reported at some point over the last decade of data (2005 - 2014), thus validating the models and only selected "current" shared species.**

@@

Finally, to have a more robust result and not categorize a species as transboundary based on its presence in a single $0.5^{\circ}$ x $0.5^{\circ}$ grid cell within an EEZ, we computed an *Area Index*. The *Area Index* consist in the proportion of a given species' overarching shared distribution between neighboring EEZs accounted for by each individual EEZ. We classified a species as transboundary if both neighboring EEZs enclosed over 25% of the species joint shared distribution. **While a species that has less than the selected threshold is not considered transboundary in this paper**, this threshold can be lowered for a more relaxed result or increased for a more conservative estimate (Fig. S3).

@@

We estimated the economic contribution in **2019** real USD of transboundary species for each country using global ex-vessel price data [@Tai2017]. The database we draw from includes ex-vessel price derived from multiple sources and a structured interpolation method (e.g. similar countries, species) to fill in data gaps [@Sumaila:2015uc]. **The dataset is harmonized with the SAU catch data to estimate yearly fishing revenue (as ex-vessel price) for all species and EEZs of this study (http://www.seaaroundus.org/data/#/feru). We report the average fishing revenue of 2005 to 2014 derived from the fishing activity within the global EEZs. Alike the catch data, for this study we did not include revenue from areas beyond national jurisdiction. We used the US Gross Domestic Product Implicit Price Deflator (average of all 2019 quarters) according to the US Saint Louis Federal Reserve Bank (https://fred.stlouisfed.org/searchresults?st=GDP+Implicit+price+deflator) to transform the original 2010 real USD value to 2019 real USD.**

We used the catch data to determine the **catch trend** of each species within an EEZ. Although this method has previously been used to estimate stock status [@Grainger:1996tk], the categories presented here are intended as catch trends, and not fishing status for each species as many environmental and social-economic factors (e.g. temperature, markets, fishing policies, **and fishing effort) affect fishing catch volume **[@Pauly:2013gs; @Branch:2008kk]. We only assessed species within each EEZ for which at least ten years of data were available between 1951 and 2014 and with at least five consecutive years of data. Three final categories were drawn for each species depending on the catch volume within each EEZ (e.g. present, maximum and minimum EEZ's historical catch) and the year (e.g. year of maximum historical catch of the species within that EEZ) (Table 2) [@Kleisner:2011wn]. Following this concept, Category A would represent fisheries that are **increasing in catch, Category B would be species that have a constant catch rate, and Category C would be species that have seen a decrease in catch in the last 10 years. Finally, we reported the predominant category over the time period considered.**

@@

Categories <- c("A - Growing Fishery",
                  "B - Stable Fishery",
                  "C - Shrinking Fishery",
                  "No Status")

@@

![Figure S1. Number of transboundary species per EEZ and their contributions to countries’ EEZ catch. The number of species is displayed in the EEZ while catch (in thousand tons) is displayed in the land polygon.](../Figures/FigS1.png)

The number of grid cells in which a species is present (*Area Index*) is expected to directly influence whether or not a species is considered transboundary. Therefore, we tested the sensitivity of our results to the *Area Index* by estimating the number of transboundary species along a gradient of values (Fig. S2). As expected, using a more relaxed value (e.g. the species is present within 10% or more of the EEZ) will result in a greater number of transboundary species (median of 40 EEZs shared per species), while a more restrictive value (e.g., 50%) results in a median estimate of 4. The current analysis utilized a 25% *Area Index* threshold (Fig. S2 - darker histogram).

@@

![Figure S2. Histogram of number of transboundary species using different Area Index threshold values. The darker histogram represents the threshold (25%) used in the present analysis](../Figures/FigS2.png)
